---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Here the data is loaded, making sure that the date column is automatically read by R as dates in the format YYYY-MM-DD. The Steps and Interval columns are classed as numeric values.
```{r data, echo=T}
library(readr)
activity <- read_csv("~/GitHub/RepData_PeerAssessment1/activity.zip", 
    col_types = cols(date = col_date(format = "%Y-%m-%d"), 
        interval = col_number(), steps = col_number()), 
    na = "NA")
View(activity)
```


## What is mean total number of steps taken per day?
```{r steps, echo=T}
#Calcualte the total number of steps each day
SumStepsDay<-aggregate(activity$steps, by=list(Category=activity$date), FUN=sum)

#Remove rows with NA values
SumStepsDay<-na.omit(SumStepsDay)

#create a histogram plot of the total number of steps each day.
hist(SumStepsDay$x)

#calculate the mean and median number of steps each day.
meanSteps<-mean(SumStepsDay$x)
medianSteps<-median(SumStepsDay$x)

```

ANSWER: The mean number of steps is `r meanSteps` and the median number of steps is `r medianSteps`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
